---
title: "p8105_hw2_yc4018"
author: "Yuxuan Chen"
date: "10/8/2021"
output: github_document
---
**Problem1**
```{r message = FALSE}
library(readxl)
library(tidyverse)
```

```{r warning = FALSE, collapse = TRUE}
trashwheel_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N535") %>% #Read the excel file
  janitor::clean_names() %>% #clean variables' names
  transform(dumpster = as.numeric(dumpster)) %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls)) #round the number of sports balls to the nearest integer

knitr::kable(trashwheel_df[0:10,])
```

```{r collapse = TRUE}
precipitation2018_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% #clean variables' names
  drop_na(month) %>%  #omit rows without precipitation data
  mutate(year = "2018") #add an variable for year
knitr::kable(precipitation2018_df)

precipitation2019_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% #clean variables' names
  drop_na(month) %>%  #omit rows without precipitation data
  mutate(year = "2019") #add an variable for year
knitr::kable(precipitation2019_df)

precipitation_df = 
  bind_rows(precipitation2018_df, precipitation2019_df) %>% 
  mutate(month = month.name[month])
knitr::kable(precipitation_df)

```

inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2019?

From the Mr. Trash Wheel dataset, trashwheel_df, there are total `r nrow(trashwheel_df)` obsesrvations. For example, 
From the combined precipitation dataset, precipitation_df, there are `r nrow(precipitation_df)` observations.


**Problem2**

I found there are five values of "2" for observations in "prez_gop", which are assumed to be typographical error and the "2" means "1". Hence, whenever there are "0" value in prez_dem imply the president was not democratic, but republican.
```{r message=FALSE, collapse=TRUE}
polsmonth_df = 
  read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>% #clean the data
  separate(mon, into = c("year", "month", "day")) %>% #break up one variable into three variables;
  transform(
    year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>%  #convert these variables to integer variables
  mutate(month = month.name[month]) %>% #replace month number with month name
  mutate(president = recode(prez_dem, '1' = "dem", '0' = "gop")) %>% #create president variable taking values gop and dem
  select(-c("prez_dem", "prez_gop","day")) #remove prez_dem, prez_gop, day variable
knitr::kable(polsmonth_df[0:10,])
```

